Identifying companies and financial actors exposed to marine tipping points

### 2024-02-27 Notes.Rmd
update with new data
marine tipping draft

>>>>

**240220**

- All data from Carmine is now added to the database. Bianca manually annotated vessels for whom we could not find orbis ID automatically (N=411). This added observations to our GUO, revenues and shareholders. However, from Orbis although the company exist, there is still missing 374 companies shareholder data, 286 companies revenues data, and 143 companies GUOs data. All data is now combined and saved in `data/cleaned_data_231214.Rda`

**240216**

- A sample of ~10k mine names were downloaded from Orbis over Xmas break. That took ~10days of computing time. The problem is many mines are not easily matchable with records in Orbis. For a few of them one can be sure due to the economic activity field in Orbis (but not always recorded), or because the mine name contains the product they exploit: diamonds, gold, etc. However, for many of them we don't know if the company info downloaded is actually a mine.
- Idea: one could circumvent this by hard coding the verification: it is mine if the econ activity is recorded, or if the name contains the metal. Other options are if the company is based on the country where the coordinates from Maus are recorded... buiding up a confidence score or something similar. Once solved, we can go ahead and download more. Bear in mind it takes long time to do so.
- Marine paper: I need to download the missing Carmine companies and then merge with complete dataset, so scripts 15 and 16 with missing data corrected by Bianca.

**231213**

- Discussion? stranded assets - link to literature and who is exposed to them
- High net worth individuals: merging the marine data, EID, fossil fuels, mining.

### 2024-02-27 marine_tipping.Rmd
marine tipping draft

Then I used Lloyds (thanks Frida!) to identify owner companies. Out of the 16 878 vessels, I got companies for 2183 of them. Using Orbis I found 793 shareholders for some of the companies. Already now you can see part of our problem: each step we try to go on the chain (vessel - owner - shareholder) we lose observations (one order of magnitude), thus to what extend are results robust? Let's check them first.

@@

```{r}
load("data/cleaned_data_231023.Rda")
load("data/boats_owners.Rda")
load("data/marine_shareholders.RData")
## some initial visualizations
countries <- readr::read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()
## correct Namibia
countries[countries$short_name == "Namibia","x2_alpha_code"] <- "NA"

@@

```{r fig.width= 5, fig.height=3, fig.align='center'}
df_boats |> 
    ggplot(aes(lli_vessel_type)) +
    geom_bar() +
    coord_flip()
```

@@

```{r}
options(DT.options = list(pageLength = 25))
datatable(guos)
```

@@

@@

```{r fig.width=5, fig.height=4}
bip <- shrs |> 
    select(name, company, ownership_total) |> 
    filter(!is.na(name)) |> 
    slice(-338) |>  # problems: the name is the IMO and the shareholder name the same
    network(directed = FALSE, bipartite = TRUE, matrix.type = "edgelist", 
            ignore.eval=FALSE, loops = TRUE)

@@

ggplot(ggnetwork::ggnetwork(bip, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = ownership_total), size = 0.35, alpha = 1
               #arrow = arrow(length = unit(2, "pt"), type = "closed")
               ) +
    geom_nodes( color = "orange", alpha = 0.4) +
    labs(title = "Bipartite network",
         subtitle = "Shareholders and companies", 
         caption = "Data source: Orbis") + 
    scico::scale_color_scico(
        "Total ownership", palette = "berlin",
        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2, "mm")))+
    theme_blank(base_size = 10) +
    theme(legend.position = "bottom")
```

The bipartite network of shareholders connected to companies if we don't have missing data. Please note that we started with over 2700 companies, but not for all there are GUOs, we only have `r shrs |>  filter(!is.na(name)) |> pull(name) |> length()` shareholder - company pairings (links), one of them is a self-link. In other words, I'm not showing the isolates, companies that we identified with Lloyds but for which Orbis did not provided shareholders. The data I'm using to build the network does not distinguish ownership over time, nor name changing. The density of the bipartite network is `r round(network.density(bip),4)`, meaning out of all possible links on the network < 0.1% are realized.

The bipartite network has two one mode projections: a network of shareholders connected to each other if they share investments to the same fishing companies (left), and a network of fishing companies connected to each other if they are invested by the same shareholders (right). As you can see, these networks are also very sparse with very few connections and no "gigant key stone actors" emerging from the data.

@@

a <- shrs |> 
    ggplot(aes(ownership_total)) +
    geom_density() +
    labs(x = "Total ownership (%)", tag = "a") +
    #scale_x_continuous(labels = scales::label_percent()) +
    theme_light()

b <- shrs |> 
    group_by(company) |> 
    summarize(n = n()) |> 
    ggplot(aes(n)) + 
    geom_bar() +
    scale_y_log10() +
    labs(x = 'Number of shareholders', y = "Companies", tag = "b") +
    theme_light()

@@

c <- ggplot(ggnetwork::ggnetwork(shr_net, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = w), size = 0.15, alpha = .5
               #arrow = arrow(length = unit(2, "pt"), type = "closed")
               ) +
    geom_nodes( color = "orange", alpha = 0.4) +
    scale_size(name = "Outdegree", range = c(0.2,4)) +
    labs(title = "Network of shareholders",
         subtitle = "Shareholders are connected if they have shares on the same fishing companies", 
         caption = "Data source: Orbis") + 
    scale_color_continuous(
        "Companies", breaks = c(1,2,3), 
        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2, "mm")))+
    theme_blank(base_size = 10) +
    theme(legend.position = "bottom")

@@

d <- ggplot(ggnetwork::ggnetwork(net, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = w), alpha = 1,
               arrow = arrow(length = unit(2, "pt"), type = "closed")) +
    geom_nodes(aes(size = loops), color = "orange", alpha = 1) +
    geom_text(aes(label = label), size = 2, color = "white") +
    scale_color_scico(
        "Shareholders",palette = "berlin", 
        guide = guide_colorbar(barwidth = unit(4,"cm"), barheight = unit(2, "mm"))) +
    scale_size(range = c(4,10), breaks = c(0, 100, 200)) +
    theme_void() + labs(tag = "d") +
    theme(legend.position = "bottom")

@@

f <- df_stats |> 
    mutate(label = case_when(
        indegree > 5 | outdegree > 5 ~ country
    )) |> 
    ggplot(aes(indegree, outdegree)) +
    geom_point(aes(size = loops), alpha = 0.5, show.legend = FALSE) +
    geom_abline(slope = 1, color = "orange", alpha = 0.5) +
    geom_text(
        aes(label = label), nudge_y = -0.5, size = 2) +
    scale_size(range = c(1,10), breaks = c(0, 100, 200)) +
    theme_light() + labs(tag = "c") +
    theme(legend.position = "bottom")

@@

@@

Second, most companies have only a few shareholders, just around 10 actually have 5 shareholders or more. Several of the investments are by shareholders on companies in the same countries, on the network these are loops. They are hard to see in the visualization but I scaled node size to correspond to these loops. Norway is the larger country with loops (> 200 shareholders invest in fishing companies inside Norway). For countries where there is actually foreign investments, Spain, UK and France are hot spots of shareholders, while Russia is a hot spot for receiving investments (see panel C).

@@

@@

1. How do you think we can improve the analysis?
2. What other data sets and questions can we ask?

In previous discussions we thought of using Eikon to get info on debt, who fund companies? In a preliminary analysis with Peruvian vessels and companies we did not find anything in Eikon. Given the sparsity of the current dataset, I doubt we will find much, but happy to discuss.

>>>>

```{r}
load("data/cleaned_data_231214.Rda")
load("data/boats_owners.Rda")
load("data/marine_shareholders.RData")
## some initial visualizations
countries <- readr::read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()
## correct Namibia
countries[countries$short_name == "Namibia","x2_alpha_code"] <- "NA"

Then I used Lloyds (thanks Frida!) to identify owner companies. Out of the 16 878 vessels, I got `r length(guos)` owners, some of them individual people, others companies. Using Orbis I found `r shrs$company |> unique() |> length()` companies and `r shrs$name |> unique() |> length()` unique shareholders with `r nrow(shrs)` investment links to the companies. This is after playing around with different strategies to complement the dataset including the data published by Carmine.

@@

@@

```{r fig.width= 5, fig.height=3, fig.align='center'}
df_boats |> 
    ggplot(aes(lli_vessel_type)) +
    geom_bar() +
    coord_flip() +
    labs(x = "Vessel type", y = "Count", caption = "Data source: Lloyds")
```

@@

```{r}
options(DT.options = list(pageLength = 10))
datatable(guos)
```

@@

# ggsave(filename = "marine_stats.png", path = "figures/",
#        plot = a + b + c & theme_light(base_size = 6),
#        device = "png", width = 7, height = 3)
```

@@

```{r fig.width=5, fig.height=4}
bip <- shrs |> 
    select(name, company, ownership_total) |> 
    filter(!is.na(name)) |> 
    filter(name != company) |>  
    unique() |> 
    # problems: the name is the IMO and the shareholder name the same
    network(directed = FALSE, bipartite = TRUE, matrix.type = "edgelist", 
            ignore.eval=FALSE, loops = TRUE)

@@

ggplot(ggnetwork::ggnetwork(bip, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = ownership_total), linewidth = 0.35, alpha = 1
               #arrow = arrow(length = unit(2, "pt"), type = "closed")
               ) +
    geom_nodes( color = "orange", alpha = 0.4) +
    labs(title = "Bipartite network",
         subtitle = "Shareholders and companies", 
         caption = "Data source: Orbis") + 
    scico::scale_color_scico(
        "Total ownership", palette = "berlin",
        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2, "mm")))+
    theme_blank(base_size = 10) +
    theme(legend.position = "bottom")
```

The bipartite network of shareholders connected to companies if we don't have missing data. Please note that we started with over 2700 companies, but not for all there are GUOs, we only have `r shrs |>  filter(!is.na(name)) |> pull(name) |> length()` shareholder - company pairings (links), one of them is a self-link. In other words, I'm not showing the isolates, companies that we identified with Lloyds but for which Orbis did not provided shareholders. The data I'm using to build the network does not distinguish ownership over time, nor name changing. The density of the bipartite network is `r network.density(bip)`, meaning out of all possible links on the network < 0.001% are realized.

The bipartite network has two one mode projections: a network of shareholders connected to each other if they share investments to the same fishing companies (left), and a network of fishing companies connected to each other if they are invested by the same shareholders (right). As you can see, these networks are also very sparse with very few connections and no "giant key stone actors" emerging from the data.

@@

a <- shrs |> 
    ggplot(aes(ownership_total)) +
    geom_density() +
    labs(x = "Total ownership (%)", tag = "a") +
    #scale_x_continuous(labels = scales::label_percent()) +
    theme_light(base_size = 8)

b <- shrs |> 
    group_by(company) |> 
    summarize(n = n()) |> 
    ggplot(aes(n)) + 
    geom_bar() +
    scale_y_log10() +
    labs(x = 'Number of shareholders', y = "Companies", tag = "b") +
    theme_light(base_size = 8)

@@

c <- ggplot(ggnetwork::ggnetwork(shr_net, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = w), linewidth = 0.15, alpha = .5
               #arrow = arrow(length = unit(2, "pt"), type = "closed")
               ) +
    geom_nodes( color = "orange", alpha = 0.4) +
    scale_size(name = "Outdegree", range = c(0.2,4)) +
    labs(title = "Network of shareholders",
         subtitle = "Shareholders are connected if they have shares on the same fishing companies", 
         caption = "Data source: Orbis") + 
    scale_color_continuous(
        "Companies", breaks = c(1,2,3), 
        guide = guide_colorbar(barwidth = unit(2,"cm"), barheight = unit(2, "mm")))+
    theme_blank(base_size = 10) +
    theme(legend.position = "bottom")

@@

d <- ggplot(ggnetwork::ggnetwork(net, arrow.gap = 0.005) ,
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(aes(color = w), alpha = 1,
               arrow = arrow(length = unit(2, "pt"), type = "closed")) +
    geom_nodes(aes(size = loops), color = "orange", alpha = 1) +
    geom_text(aes(label = label), size = 2, color = "white") +
    scale_color_scico(
        "Shareholders",palette = "berlin", 
        guide = guide_colorbar(barwidth = unit(4,"cm"), barheight = unit(2, "mm"))) +
    scale_size(range = c(4,10), breaks = c(0, 100, 200)) +
    theme_void(base_size = 7) + labs(tag = "d") +
    theme(legend.position = "bottom")

@@

f <- df_stats |> 
    mutate(label = case_when(
        indegree > 5 | outdegree > 5 ~ country
    )) |> 
    ggplot(aes(indegree, outdegree)) +
    geom_point(aes(size = loops), alpha = 0.5, show.legend = FALSE) +
    geom_abline(slope = 1, color = "orange", alpha = 0.5) +
    geom_text(
        aes(label = label), nudge_y = -0.5, size = 2) +
    scale_size(range = c(1,10), breaks = c(0, 100, 200)) +
    theme_light(base_size = 7) + labs(tag = "c") +
    theme(legend.position = "bottom")

@@

ggsave(
    path = "figures/", filename = "marine_nets.png", device = "png",
    width = 7, height = 5, dpi = 500,
    plot = ((a/b/f) | d) + plot_layout(widths = c(1,3))
)

@@

Second, most companies have only a few shareholders. Several of the investments are by shareholders on companies in the same countries, on the network these are loops. They are hard to see in the visualization but I scaled node size to correspond to these loops. Norway is the larger country with loops (> 200 shareholders invest in fishing companies inside Norway). For countries where there is actually foreign investments, Spain, UK and France are hot spots of shareholders, while Russia is a hot spot for receiving investments (see panel C).

@@

**Global Ultimate Owners**

```{r fig.width=8, fig.height=8}
load("data/cleaned_data_231214.Rda")

guos <- guos |>
    mutate(class = case_when(
        str_detect(guo, "Public") ~ "Public",
        str_detect(guo, "Private") ~ "Private",
        str_detect(guo, "Formerly") ~ "Former public",
        .default = guo
    )) |> # pull(class) |> table()
    mutate(
        ## remove all the cases
        guo = str_remove(guo, "Private The Global Ultimate Owner of this controlled subsidiary is "), 
        guo = str_remove(guo, "Formerly publicly quoted The Global Ultimate Owner of this controlled subsidiary is "),
        guo = str_remove(guo, "Private This entity is a marine vessel of "),
        guo = str_remove(guo, "Publicly quoted The Global Ultimate Owner of this controlled subsidiary is "),
        guo = str_remove(guo, "Private This entity was a branch of "),
        guo = str_remove(guo, "Private This entity is a branch of "), 
        guo = str_remove(guo, "Private This entity is a foreign company of "),
        ) |> #pull(guo) |> unique() #|> 
    mutate(
        guo = case_when(
            guo =="Private This entity is a marine vessel." ~ NA,
            guo == "Private The Global Ultimate Owner of this company could not be identified."  ~ NA,
            guo == "Private This entity is a branch." ~ NA,
            guo == "Publicly quoted This company is Independent (but not the Global Ultimate Owner of a Corporate Group)"  ~ NA,
            guo == "Publicly quoted This company is a Single location" ~ NA, 
            guo == "Private" ~ NA,
            guo == "Private This entity is a foreign company." ~ NA,
            guo == "Private This company is Independent (but not the Global Ultimate Owner of a Corporate Group)" ~NA,
            guo == "Publicly quoted This company is the Global Ultimate Owner of the corporate group"  ~ orbis_name,
            guo == "Private This company is the Global Ultimate Owner of the corporate group"  ~ orbis_name,
            guo == "Formerly publicly quoted This company is the Global Ultimate Owner of the corporate group" ~ orbis_name,
            guo == "Private This company is a Single location" ~ NA,
            .default = guo
        )
    )  |> 
    mutate(person = str_starts(guo, pattern = "MR|MRS|MS")
    )

a <- guos |> 
    ggplot(aes(class)) +
    geom_bar() +
    coord_flip()

b <- guos |> 
    ggplot(aes(person)) + 
    geom_bar() +
    coord_flip()

c <- guos |> 
    filter(!is.na(guo), person == FALSE) |> 
    group_by(guo) |> 
    summarize(n = n()) |>
    arrange(n) |> 
    slice_tail(n = 25) |> 
    mutate(guo = as_factor(guo)) |> 
    ggplot(aes(n, guo)) +
    geom_col()

(a + b) / c & theme_light(base_size = 8)

@@

1. What to do with GUOs? Do we include them on the analysis?

In previous discussions we thought of using Eikon to get info on debt, who fund companies? In a preliminary analysis with Peruvian vessels and companies we did not find anything in Eikon. Given the sparsity of the current dataset, I doubt we will find much, but happy to discuss. I also tried with the Datastream (part of Refinitiv) platform thanks to IVL access. I did not find much info on debt, the companies are too small / unknown.

### 2024-02-27 paper/manuscript.Rmd
marine tipping draft

---
title: "Identifying companies and financial actors exposed to marine tipping points"
author: |
    | \small Juan C. Rocha^1^, Jean-Baptiste Jouffray^1^, Paula A. Sánchez^1,2^, Victor Galaz^1^
    | \footnotesize ^1^Stockholm Resilience Centre, Stockholm University
    | \footnotesize ^2^Germany?
    | \small \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
  word_document:
csl: /Users/juanrocha/Documents/styles/apa.csl
#bibliography: refs.bib
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, array, tabularx, inputenc}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
- \linenumbers
- \usepackage{setspace}
---

@@

- Author order tbd
- Target journal: Nature family?, GEC, Finance journal? ~ 4000w, 4 figs, 40refs
- Word budget
    - Intro: 1000w
    - Method: 800w
    - Results: 1000w
    - Discussion: 800w
    - Conclusions: 400w

@@

@@

_Companies_: Based on the geographic exposure to regime shifts, we then identified fishing companies operating in these regions. We used permits and licences issued in Alaska by the National Oceanic and Atmospheric Administration in the United States ([NOAA](https://www.fisheries.noaa.gov/alaska/commercial-fishing/permits-and-licenses-issued-alaska)) to list companies with permits to fish Alaskan pollock, one of the most profitable resources in the region. We used fishing permits issued by the _Servicio nacional de Pesca y Acuicultura_ or the National Fishing and Aquaculture authority of Chile, to companies with the purpose of harvesting anchovies and sardines.

- check with Paula, we should have Peru as well
- what to do with tuna? we only have O

@@

@@

- Limitation: publicly listed companies, and legal fisheries.

@@

>>>>

---
title: "Identifying companies and financial actors exposed to marine tipping points"
author: |
    | \small Juan C. Rocha^1^, Jean-Baptiste Jouffray^1^, Frida Bengtsson^1^, Bianca-Ioana Voicu^1^,  Paula A. Sánchez^1,2^, Victor Galaz^1^
    | \footnotesize ^1^Stockholm Resilience Centre, Stockholm University
    | \footnotesize ^2^Germany?
    | \small \texttt{\href{mailto:juan.rocha@su.se}{\nolinkurl{juan.rocha@su.se}}}
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
  word_document:
csl: /Users/juanrocha/Documents/styles/apa.csl
bibliography: refs.bib
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, array, tabularx, inputenc}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
- \linenumbers
- \usepackage{setspace}
---

@@

- Author order tbd
- Target journal: Nature family?, SciAdv, GEC, Finance journal? ~ 4000w, 4 figs, 40refs

## Outline

- Intro: 1000w
    -  Marine ecosystems are expected to undergo large biological reorganizations
    - Exposure to regime shifts are mediated by pressures from climate change, fishing pressure, or pollution.
    - Companies and financial actors are part of the problem but can also be part of the solutions.
    - RQ: **Which financial actors are exposed to regime shift risks, and thus have the agency to avoid them or can play a role in navigating transformations?**
- Method: 800w
    - Identifying areas prone to regime shifts:
        - High probability of marine heatwaves
        - High probability or regime shifts, singals of proximity to tipping points
    - Identify vessels who fish in these areas
    - Identify companies who operate these vessels (owners). Identify shareholders who have shares on these companies. Identify global ultimate owners. 
    - Network statistical models to explore what influence the likelihood of investment.
- Results: 1000w
    - The network of ownership and shareholder investments is sparse lacking a backbone or common financial giants
    - Despite the sparsity of the ownership network, there are important flows of investments across national borders.
    - Probability of investments: ergms
- Discussion: 800w
- Conclusions: 400w

@@

- Identifying areas prone to regime shifts:
    - High probability of marine heatwaves
    - High probability or regime shifts, singals of proximity to tipping points
- Identify vessels who fish in these areas
- Identify companies who operate these vessels (owners). Identify shareholders who have shares on these companies. Identify global ultimate owners. 
- Network statistical models to explore what influence the likelihood of investment.

@@

_Companies_: Based on the geographic exposure to regime shifts, we then identified fishing companies operating in these regions.

@@

- The network of ownership and shareholder investments is sparse lacking a backbone or common financial giants (SM nets figures).
    - contrary to previous results of SeaBOS, and financial giants: why? Ownership structure: few owners who own 1/2 or the entirety of the company, many family and small scale businesses. Market conglomeration probably emerge at the level of intermediaries, but from the data, that is not the case for the captains who make decisions on where to fish.
    - Yet, most vessels are Chinesse, followed by Japan and USA. Most shareholders are based in Norway but they do invest mostly on Norwegian fishing companies.

- Despite the sparsity of the ownership network, there are important flows of investments across national borders. 
    - International investments are dominated by Spain, UK and France as countries where investments originate, while Russia is a country who disproportionately receives investments.

@@

- Contrast to other papers of key stone actors: fisheries does not seem to be highly concentrated, at least not at the level of decisions that expose to risk of regime shifts -- where to go fishing and when.

- Limitation: publicly listed companies, and legal fisheries.
    - We do not take care of timing: a company can change owners during the time period. We do not control for it because chances are the previous owners reinvest in fishing related activities, there is no way to track investments.

@@

Grants XYZ
